Software architectures to integrate workflow engines in science gateways

### 2016-10-12 answers.tex
Even more answers to the reviewers

\usepackage{hyperref}
\newenvironment{review}%          environment name
{\textbf{Reviewer comment:}\begin{quote}}% begin code
{\end{quote}}%  
\newenvironment{answer}%          environment name
{\textbf{Answer:}\begin{small}\begin{quote}}% begin code
{\end{quote}\end{small}}%

@@

\newcommand{\revised}[1]{\color{blue} #1\color{black}}

@@

\todo{just remove CSGF, Airvata and pgrade from the table, keeping the
  exponents. But we don't have examples of gateways using Airvata,
  still need to be found.}

@@

\todo{ponderation again}

@@

@@

\todo{answer metric by metric, focusing on E1 and E4. For E4 it's clear. For E1, }

@@

>>>>

\usepackage{hyperref}
\usepackage{xspace}
\newenvironment{review}%          environment name
{\textbf{Reviewer comment:}\begin{quote}}% begin code
{\end{quote}}%  
\newenvironment{answer}%          environment name
{\textbf{Answer:}\begin{small}\begin{quote}}% begin code
{\end{quote}\end{small}}%

@@

\newcommand{\revised}[1]{\color{blue} #1\color{black}\xspace}

@@

\begin{answer}
We thank the reviewer for the thorough review and useful comments.

\todo{just remove CSGF, Airvata and pgrade from the table, keeping the
  exponents. But we don't have examples of gateways using Airvata,
  still need to be found.}
\end{answer}

@@

\begin{answer}
  We assume that the reviewer is referring to the I$_1$ metric which
  is, indeed, lower for the tight integration than for any other
  architecture (it is equal for task encapsulation). We agree with the
  reviewer that the difficulty to update a component depends on the
  nature of this component. However, we believe that a quantification
  of this difficulty is very much dependent on the particular system
  under study. For instance, in some cases, tightly integrating a
  workflow engine may be straightforward (e.g. when the science
  gateway has the appropriates hooks) while developing a service may
  require a lot of work if it has to be done from scratch.

  We believe that this discussion comes down to the assignments of
  different weights to the interactions and components in the
  architectures. As discussed before, these weights can now be
  adjusted using the spreadsheet at \url{https://frama.link/LrPRsCQp}
  to better describe a particular system.
\end{answer}

@@

\begin{answer}
We thank the reviewer for the thorough review and useful comments.

Regarding componentization: our architectures are described in terms
of software components and interactions among these components (see
Figure 1). So we believe that our approach fits in the context of
component-based software engineering. The problem that we address is
to determine what components should be defined and how they should
interact to optimize the metrics. For instance, the comparison between
the tightly integrated and service invocation architectures evaluates
whether the workflow engine should be a specific component or if it
should be integrated in the science gateway. Besides, the reviewer's
comment touches upon the granularity at which components are
defined. We extended the first paragraph of Section 2 to address this issue:
\revised{Software
  components are defined at a granularity such that two components may
  potentially be operated by different teams on different hardware
  systems. For instance, the science gateway and infrastructure are
  usually distinct components. In real systems, software components
  may exist at a finer granularity, to implement functions such as
  authentication, data persistency or logging. Such fine-grained
  components are not covered by our analysis since we focus on the
  interactions between the science gateway, the infrastructure and the
  workflow engine.}

The discussion on software layers is more difficult. Roughly, our
architectures could be described as multilayered software
architectures. For instance, the tight integration has two layers
(science gateway and infrastructure), the service invocation has
three, the task encapsulation has two, etc. However, strictly
speaking, these are not layers because it would mean that layer n-1
can only communicate with layer n, which is not the case. For
instance, in service invocation, layer 0 (science gateway)
communicates with layer 2 (infrastructure) to transfer data. And in
task encapsulation, layer 1 (infrastructure) communicates with layer 0
(science gateway). Thus, we don't think that referring to multilayered
architectures is accurate here.

We don't think that our architectures lead to conclude systematically
that ``the more layers to components in the system, the worse it
scores'', as suggested by the reviewer. For instance, service
invocation scores better than tight integration in E1 and S1 due to
the additional ``layer'' used to host the workflow engine. Likewise,
nested workflows score better than tight integration for S2. However,
our metrics do imply that more layers leads to more integration
complexity and reduced robustness, which we think is very commonly
experienced in distributed systems, where  communication between
``layers'' involve wide-area network communications between machines
that may be operated by different people. We added a note in the definition of our robustness metrics to reflect this discussion:
\begin{quote}
 Robustness is measured here as a consequence of global
complexity, since complex architectures tend to be more prone to
failure, \revised{in particular in distributed systems where
  interactions between components involve wide-area network
  communications between machines that may be operated by different
  people}.
\end{quote}
\end{answer}

@@

\begin{answer}
  The issue raised by the reviewer needs to be discussed metric by
  metric rather than on the global extensibility score. The pool and
  nested architectures differ in metrics E$_1$ (adding a new engine
  type) and E$_4$ (adding a new infrastructure). For E$_4$ (adding a
  new infrastructure), it seems clear to us that the nested
  architecture is more difficult to extend because it has two workflow
  engines while the pool has only one. Therefore, interactions
  \texttt{b} and \texttt{c$_2$} have to be counted twice, which
  explains the difference in scores for E$_4$.

  The situation highlighted by the reviewer is related to E$_1$ (new
  engine type). If a new engine is added to the pool model, the
  following interactions and components have to be updated:
 \begin{enumerate}
 \item The engine has to be wrapped in the SHIWA Agent, which includes
   the reading/writing of ORE bundles so that the agent can get/pass
   them to the pool through interaction \texttt{f}. In practice, the
   agent component has to be modified to convert ORE bundles to the
   input format supported by the engine, and to assemble ORE bundles
   from the output format produced by the engine. As suggested by the
   reviewer, another implementation possibility would be to modify the
   engine to read/write ORE bundles so that the integration in the
   agent is straightforward.
 \item The engine has to be able to transfer data to the
   infrastructure used in the science gateway, i.e. interaction
   \texttt{c$_2$} has to be implemented in the engine.
 \item The engine has to be able to submit tasks to the infrastructure
   used in the science gateway, i.e. interaction \texttt{b} has to be
   implemented in the engine.
 \end{enumerate}
This is why E$_1$=3 for the pool model.

If a new child engine is added to the nested model then the following interactions an and components have to be updated:
\begin{enumerate}
\item A workflow service has to be created for the workflow engine.
\item The parent engine has to be interfaced with the new workflow
  service, i.e., the workflow service should be able to digest ORE
  bundles, or the parent engine should convert ORE bundles to the
  format supported by the workflow service. ORE bundles will be
  exchanged through interaction \texttt{d}.
 \item The child engine has to be able to transfer data to the
   infrastructure used in the science gateway, i.e. interaction
   \texttt{c$_2$} has to be implemented.
 \item The child engine has to be able to submit tasks to the infrastructure
   used in the science gateway, i.e. interaction \texttt{b} has to be
   implemented,
\end{enumerate}
This is why E$_4$=4 for the nested model.

The situation described by the reviewer focuses on item 1 of the pool
extension and item 1 and 2 of the nested workflow extension
(infrastructure is not a discriminant factor here). First, the
reviewer suggest that the availability of ORE bundles in the SHIWA
framework standardizes the communication with workflow engines so that
it becomes possible to have a generic workflow service that can
consume and produce ORE bundles regardless of the particular engine(s)
it wraps. So indeed, item 1 of the nested case disappears. This is a
typical case where, as we wrote in section 7.2, \emph{The particular
technical or historical context of a science gateway project may also
influence the \revised{evaluation} of an architecture.} In this
case, the possible availability of a generic workflow service based on
ORE bundles is contextual information that can hardly be integrated in
the generic analysis. We believe that our framework helps evaluate
the impact of technologies such as ORE bundles on architectures, as
discussed here.

Second, the
reviewer suggests that the integration of the workflow engine in the
agent (pool model) is easier than in the service because the engine
has to produce and consume ORE bundles in the former case while it
only has to consume them in the latter case. This is true again, but
it comes down to the discussion on the weights of the different
software components. In this particular case, it would make sense to
assign a higher weight to the agent than to the service, which could
be done through the public spreadsheet.  
\end{answer}

@@

\begin{answer}
  We agree with the reviewer that functionality is important. This was
  exactly the goal of the ``Specific features'' evaluation
  criterion. We renamed this criterion ``Functionality'' as we find it a much better name indeed. We also renamed O$_1$ to F$_1$ and O$_2$ to F$_2$.
\end{answer}
